---
title: "Session 2, part IV: Exercise - Building a simple neural network using Keras and Tensorflow"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Example of Keras workflow

### 1. Prepare Data
```{r, eval=FALSE}
# Prepare data
# Scale features
# Define
x_train
y_train
x_test
y_test
```

(Note, splitting the data into a `train` and a `test` set is called cross validation, more on this later)

### 2. Define Architecture

We can define the architecture of an ANN with `4` input neurons, `4` neurons in the first hidden layer, `3` neurons in the second hidden layer and `3` output neurons like so:

```{r, eval=FALSE}
model = keras_model_sequential()
model %>% 
  layer_dense(units = 4, activation = 'relu', input_shape = 4) %>% 
  layer_dense(units = 3, activation = 'softmax')
model %>% summary
```

This gives a total of 35 parameters to be fitted: 4 for each input neuron connected to the hidden layer, plus an additional 4 for the associated first bias neuron and 3 for each of the hidden neurons connected to the output layer, plus an additional 3 for the associated second bias neuron. I.e. `(4 * 4) + 4 + (4 * 3) + 3 = 35`

### 3. Compile Model

```{r, eval=FALSE}
model %>% compile(
  loss      = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics   = c('accuracy')
)
```

### 4. Train the Model

```{r, eval=FALSE}
history = model %>% fit(
  x = x_train, y = y_train,
  epochs           = 200,
  batch_size       = 20,
  validation_split = 0
)
```

### 5. Evaluate Model Performance
```{r, eval=FALSE}
perf = model %>% evaluate(x_test, y_test)
```

## Summary

1. Prepare Data
2. Define Architecture
3. Compile Model
4. Train the Model
5. Evaluate Model Performance
6. Adjust hyperparameters and redo steps 1-5

## Exercise: Keras TensorFlow on Iris

1. Goto [Building a simple neural network using Keras and Tensorflow
](https://github.com/leonjessen/keras_tensorflow_on_iris)
2. Copy the code to the editor of your choice
3. Make sure the script can run
4. Try running the script a few times without changing anything, do you get consistent results?
5. Play around with hyperparameters and architecture
